\mytitle

### 2023-05-13 chapters/concl.tex
Remaining emina edits

There is still more work to be done in support of this thesis.
% TODO better transition sentence
%In addition to the cases presented, there are many more classes of systems
%that could benefit from program synthesis tools.
One interesting direction for work in this area is expanding on \jitsynth
to allow for some class of compiler optimizations.
In order to further improve the runtime performance of synthesized compilers,
\jitsynth could use superoptimization techniques to minimize the size
of discovered minicompilers.
Exploring how to best apply these techniques to improve \jitsynth's effectiveness
would be an interesting direction for future work.
%One interesting direction for work in this area is
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
%and properties for staking protocols are specified in a fairly straightforward mannor,
%but complex required invariants on ___ make implementing staking mechanisms difficult.
%Automating this implementation step could prevent bugs in these protocols
%and the massive loss of funds that comes with these exploits.
Another promising direction lies in further improving \depsynth's usability.
Existing techniques for litmus test generation and partial interpretation of litmus tests
\cite{mohan:crashmonkey,bornholt:memsynth}
could allow \depsynth to synthesize dependency rules more quickly.
% Distributed protocol synthesis: https://www.cis.upenn.edu/~alur/udupa-thesis.pdf
Finally, my overall thesis would be further supported by a larger
body of synthesis tools for a wide array of modern systems.
\cref{c:depsynth} discusses existing works in synthesis for
techniques for both memory models and distributed protocols,
but other promising system domains for program synthesis %application
include blockchain staking protocols and security monitors for IFC kernels.
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
Security and liveness properties for staking protocols are specified in a fairly straightforward manner,
but complex invariants on the blockchain state make implementing staking mechanisms difficult
\cite{fahad:pos,wenting:secure-pos,aggelos:verified-pos}.
Automating this implementation step could prevent bugs in staking protocols
and the massive loss of funds that comes with these exploits.
Information flow control (IFC) kernels give developers the ability to enforce
a variety of security properties by limiting the information flow between applications.
However, correctly implementing these restrictions is prohibitively difficult.
Existing work enables push-button verification of these security properties for IFC kernels \cite{sigurbjarnarson:nickel},
but still requires a manually-written implementation.
By synthesizing security monitors for IFC kernels that guarantee these properties,
the high implementation burden for IFC security could be removed entirely.
Overall, program synthesis tools can help systems developers build correct, performant software
more effectively and easily for a wide spectrum of system domains.

>>>>

There are several interesting directions for extending the work presented in this thesis.
%There is still more work to be done in support of this thesis.
% TODO better transition sentence
%In addition to the cases presented, there are many more classes of systems
%that could benefit from program synthesis tools.
One is expanding on \jitsynth
to allow for some class of compiler optimizations.
In order to further improve the runtime performance of synthesized compilers,
\jitsynth could use superoptimization techniques to minimize the size
of discovered minicompilers.
Exploring how to best apply these techniques to improve \jitsynth's effectiveness
is a promising direction for future work.
%One interesting direction for work in this area is
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
%and properties for staking protocols are specified in a fairly straightforward mannor,
%but complex required invariants on ___ make implementing staking mechanisms difficult.
%Automating this implementation step could prevent bugs in these protocols
%and the massive loss of funds that comes with these exploits.
Another interesting direction lies in further improving \depsynth's usability.
Existing techniques for litmus test generation and partial interpretation of litmus tests
\cite{mohan:crashmonkey,bornholt:memsynth}
could allow \depsynth to synthesize dependency rules more quickly.
% Distributed protocol synthesis: https://www.cis.upenn.edu/~alur/udupa-thesis.pdf
Finally, my overall thesis would be further supported by a larger
body of synthesis tools for a wide array of modern systems.
\cref{c:depsynth} discusses existing works in synthesis for
techniques for both memory models and distributed protocols,
but other promising system domains for program synthesis %application
include blockchain staking protocols and security monitors for IFC kernels.
%automating the implementation of blockchain staking protocols with program synthesis.
%Existing work has developed techniques for synthesizing more traditional distributed protocols, \todo{cite transit}
Security and liveness properties for staking protocols are specified in a fairly straightforward manner,
but complex invariants on the blockchain state make implementing staking mechanisms difficult
\cite{fahad:pos,wenting:secure-pos,aggelos:verified-pos}.
Automating this implementation step could prevent bugs in staking protocols
and the massive loss of funds that comes with these exploits.
Information flow control (IFC) kernels give developers the ability to enforce
a variety of security properties by limiting the information flow between applications.
However, correctly implementing these restrictions is prohibitively difficult.
Existing work enables push-button verification of these security properties for IFC kernels \cite{sigurbjarnarson:nickel},
but still requires a manually-written implementation.
By synthesizing security monitors for IFC kernels that guarantee these properties,
the high implementation burden for IFC security could be removed entirely.
Overall, program synthesis tools can help systems developers build correct, performant software
more effectively and easily for a wide spectrum of system domains.

### 2023-05-13 chapters/depsynth/impl.tex
Remaining emina edits

\subsection{Attempting graph search with a \tord solver.}
One bottleneck \depsynth
comes from the time it takes to search for a minimal
happens-before graph for a given litmus test.
This section describes a strategy that we explored to limit this bottleneck.
Specifically, we encode the happens-before graph search problem
in the theory of ordering consistency ($\tord$) \cite{n:tord}
and execute the search with an existing $\tord$ solver.
Ultimately, we found that this different search strategy did not 
improve the overall runtime, but we find this negative
result worth reporting.

@@

In an attempt to speed up the overall search time for total orders over writes,
we encode the problem in $\tord$ and offloading to a $\tord$ solver.
$\tord$ defines predicates for describing partial-ordering constraints
on reads and writes for multi-threaded programs, % TODO
but can also be used to describe orderings of storage system writes.
Specifically, $\tord$ introduces the predicate $\prec$, where
$ w_i \prec w_j $ encodes the fact that write $w_i$ occured before
write $w_j$.
To encode our search in $\tord$,
we describe the existence of each happens-before graph edge
between write events $e_i$ and $e_j$ as the boolean variable $d_{ij}$,
then introduce the following constraint: $ d_{ij} \rightarrow e_i \prec e_j $.
By solving these constraints along with the user-defined constraints specifying
crash consistency for the storage system, we arrive at a total order over all
writes for the given litmus test.

Existing work defines a DPLL(T)-style solver for $\tord$ \cite{n:tord},
which we use for solving the constraints specified above and
determining a total order over writes for single litmus tests.
However, by doing so, we empirically found that the solver
slightly diminished the performance of \depsynth for all three of our evaluated case studies
(in the case of \shardstore, slowing the rules search from
\shardstoresynthesistime minutes to \shardstoresynthesistimetord minutes).
Intuitively, we believe this performance degredation is an artifact of 
bad initial assignments from the SAT solver
and a resulting a large number of iterations in the DPLL(T) loop.
For storage systems with increasingly complex behavior,
\depsynth may require much larger litmus tests,
which may make the benefits of searching for orders with $\tord$ more apparent.
However, for the case studies considered in this paper,
we found our original search algorithm from \cref{sec:alg:onetest}
to outperform the $\tord$-directed search.

\subsection{Faster Consistency Checks for Litmus Tests}
\label{s:improved-search}
As \cref{sec:eval:shardstore} discusses,
a large majority of the litmus tests seen during the \depsynth
algorithm do not result in new dependency rules
(\depsynth generated new rules for only \shardstoretestsused of \shardstoreinputtests
 litmus tests seen for \shardstore).
However, for each litmus test in the input set, \depsynth
must perform a check that the accumulated dependency rules
guarantee the crash consistency of the test.
For synthesizing rules in \shardstore,
these take up \shardstoreverificationminutes minutes (\shardstoreverificationpercent\%)
of the overall search time.

@@

Experimentally, we observed that for these queries with long runtimes,
decomposing the query by concretizing one or more write-committed variables
drastically speeds up solving time.
(averaging $\splitvariableavgspeedup \times$ and at most $\splitvariablebestspeedup \times$ speedup for a single litmus test).
Specifically, considering the set of write-committed variables $V$,
the happens-before graph $G\subset V\times V$,
and the original encoding for the litmus test crash consistency check $P(G)$,
we choose a variable $v_0\in V$ and issue the following two queries:
\begin{enumerate}
  \item $v_0 \wedge \forall v\in V. (v, v_0)\in G \rightarrow v) \wedge P(G)$
  \item $\neg v_0 \wedge \forall v\in V. (v, v_0)\in G \rightarrow \neg v) \wedge P(G)$
\end{enumerate}
The litmus test is crash consistent with respect to the dependency rules
if and only if both queries succeed.
This concretization can be repeated for any number of variables,
resulting in $2^n$ queries for $n$ concretized variables.
Ultimately, this optimization gives a speedup of $\splitvariableoverallspeedup \times$
for the \shardstore case study discussed in \cref{sec:eval:shardstore},
reducing end-to-end runtime from \shardstoresynthesistime minutes to \splitvariablenewtime minutes.

>>>>

\subsection{Attempting graph search with a \tord solver.}
One bottleneck in \depsynth
comes from the time it takes to search for a minimal
happens-before graph for a given litmus test.
This section describes a strategy that we explored to limit this bottleneck.
Specifically, we encode the happens-before graph search problem
in the theory of ordering consistency ($\tord$) \cite{n:tord}
and execute the search with an existing $\tord$ solver.
Ultimately, we found that this different search strategy did not 
improve the overall runtime, but we find this negative
result worth reporting.

@@

In an attempt to speed up the overall search time for total orders over writes,
we encode the problem in $\tord$ and offload to a $\tord$ solver.
$\tord$ defines predicates for describing partial-ordering constraints
on reads and writes for multi-threaded programs, % TODO
but can also be used to describe orderings of storage system writes.
Specifically, $\tord$ introduces the predicate $\prec$, where
$ w_i \prec w_j $ encodes the fact that write $w_i$ occured before
write $w_j$.
To encode our search in $\tord$,
we describe the existence of each happens-before graph edge
between write events $e_i$ and $e_j$ as the boolean variable $d_{ij}$,
then introduce the following constraint: $ d_{ij} \rightarrow e_i \prec e_j $.
By solving these constraints along with the user-defined constraints specifying
crash consistency for the storage system, we arrive at a total order over all
writes for the given litmus test.

Existing work defines a DPLL(T)-style solver for $\tord$ \cite{n:tord},
which we use for solving the constraints specified above and
determining a total order over writes for single litmus tests.
However, by doing so, we empirically found that the solver
slightly diminished the performance of \depsynth for all three of our evaluated case studies
(in the case of \shardstore, slowing the rules search from
\shardstoresynthesistime to \shardstoresynthesistimetord).
Intuitively, we believe this performance degredation is an artifact of 
bad initial assignments from the SAT solver
and a resulting a large number of iterations in the DPLL(T) loop.
For storage systems with increasingly complex behavior,
\depsynth may require much larger litmus tests,
which may make the benefits of searching for orders with $\tord$ more apparent.
However, for the case studies considered in this paper,
we found our original search algorithm from \cref{sec:alg:onetest}
to outperform the $\tord$-directed search.

\subsection{Faster Consistency Checks for Litmus Tests}
\label{s:improved-search}
As \cref{sec:eval:shardstore} discusses,
a large majority of the litmus tests seen during the \depsynth
algorithm do not result in new dependency rules
(\depsynth generated new rules for only \shardstoretestsused of \shardstoreinputtests
 litmus tests seen for \shardstore).
However, for each litmus test in the input set, \depsynth
must perform a check that the accumulated dependency rules
guarantee the crash consistency of the test.
For synthesizing rules in \shardstore,
these take up \shardstoreverificationminutes (\shardstoreverificationpercent)
of the overall search time.

@@

Experimentally, we observed that for these queries with long runtimes,
decomposing the query by concretizing one or more write-committed variables
drastically speeds up solving time.
(averaging $\splitvariableavgspeedup \times$ and at most $\splitvariablebestspeedup \times$ speedup for a single litmus test).
Specifically, considering the set of write-committed variables $V$,
the happens-before graph $G\subset V\times V$,
and the original encoding for the litmus test crash consistency check $P(G)$,
we choose a variable $v_0\in V$ and issue the following two queries:
\begin{enumerate}
  \item $v_0 \wedge \forall v\in V. (v, v_0)\in G \rightarrow v) \wedge P(G)$
  \item $\neg v_0 \wedge \forall v\in V. (v, v_0)\in G \rightarrow \neg v) \wedge P(G)$
\end{enumerate}
The litmus test is crash consistent with respect to the dependency rules
if and only if both queries succeed.
This concretization can be repeated for any number of variables,
resulting in $2^n$ queries for $n$ concretized variables.
Ultimately, this optimization gives a speedup of $\splitvariableoverallspeedup \times$
for the \shardstore case study discussed in \cref{sec:eval:shardstore},
reducing end-to-end runtime from \shardstoresynthesistime to \splitvariablenewtime.

### 2023-05-13 chapters/jitsynth/impl.tex
Fix bib/refs
Remaining emina edits

\section{Implementation}\label{s:impl}
We implemented \jitsynth as described in \autoref{s:overview} using
Rosette~\cite{torlak:rosette} as our host language.
%Without bounding the length of the compiler, the space explored by \jitsynth is infinite, so we use a timeout
%of 48 hours to bound execution.
Since the search spaces for different compiler lengths are disjoint,
the \jitsynth implementation searches these spaces in parallel~\cite{bornholt:synapse}.
%In addition, \jitsynth implements several optimizations
%in order to synthesize real-world JIT compilers, as follows.
%
\if 0
\paragraph{Synthesizing Common Sequences.}
In addition to separately synthesizing pre-loads for parameters,
\jitsynth also synthesizes a few other small sequences used commonly in compiled programs.

@@

\paragraph{PC Equivalence.}
To determine the equivalence function between the source and target PC,
\jitsynth considers both the user-input PC multiple $m_{\PC}$,
as well as the length of each \minicompiler $L$.
Specifically, we say that source $\PC_S$ and target $\PC_T$
are equivalent when $\PC_S \times L \times m_{\PC} = \PC_T$.
To avoid arithmetic overflow,
\jitsynth bounds the maximum value of the $\PC_S$ to
$\lfloor\frac{2^B}{L \times m_{\PC}}\rfloor$,
where $B$ is the bitlength of the source PC.

\paragraph{NOP Padding and Removal.}
\jitsynth synthesizes \minicompilers of equal length to ensure that
the target PC for jumps can be computed as a multiple of the source PC.
%
To do so, \jitsynth pads instruction sequences with NOPs to match the length
of the largest \minicompiler.
%
To mitigate the performance impact this incurs, we implemented a trusted compiler
pass that removes the NOP instructions while preserving the correctness of
the compiled code.
\fi
%
%\paragraph{ARM Fuel Function.}
%For simplicity,
We use $\Phi(\vec{p}) = \texttt{length}(\vec{p})$ as the fuel function for all
languages studied in this paper. This provides sufficient fuel for evaluating
programs in these languages that are accepted by the OS kernel. For example, the
Linux kernel requires eBPF programs to be loop-free, and it enforces this
restriction with a conservative static check; programs that fail the check are
not passed to the JIT~\cite{gershuni:crab-ebpf}.
\tighten

>>>>

\section{Implementation}\label{s:impl}
We implemented \jitsynth as described in \autoref{s:overview} using
Rosette~\cite{torlak:rosette} as our host language.
%Without bounding the length of the compiler, the space explored by \jitsynth is infinite, so we use a timeout
%of 48 hours to bound execution.
Since the search spaces for different compiler lengths are disjoint,
the \jitsynth implementation searches these spaces in parallel~\cite{bornholt:synapse}.
%In addition, \jitsynth implements several optimizations
%in order to synthesize real-world JIT compilers, as follows.
%
%
%\paragraph{ARM Fuel Function.}
%For simplicity,
We use $\Phi(\vec{p}) = \texttt{length}(\vec{p})$ as the fuel function for all
languages studied in this paper. This provides sufficient fuel for evaluating
programs in these languages that are accepted by the OS kernel. For example, the
Linux kernel requires eBPF programs to be loop-free, and it enforces this
restriction with a conservative static check; programs that fail the check are
not passed to the JIT~\cite{gershuni:crab-ebpf}.

\paragraph{Synthesizing Common Sequences.}
In addition to separately synthesizing pre-loads for parameters,
\jitsynth also synthesizes a few other small sequences used commonly in compiled programs.

@@

\paragraph{NOP Padding and Removal.}
\jitsynth synthesizes \minicompilers of equal length to ensure that
the target PC for jumps can be computed as a multiple of the source PC.
%
To do so, \jitsynth pads instruction sequences with NOPs to match the length
of the largest \minicompiler.
%
To mitigate the performance impact this incurs, we implemented a trusted compiler
pass that removes the NOP instructions while preserving the correctness of
the compiled code.
\tighten
